Gapfilling the number of farms to add for tuna countries based on tonnes of production

### 2021-08-12 marine/last_step_rasterize/shrimp_rasterize_production.Rmd
cleaned up scripts and READMEs

all_farms <- readRDS("data/all_modeled_marine_farms.rds")

shrimp_farms <- all_farms %>%
  filter(type == "shrimp")

@@

# CHN:HKG
# CHN:MAC
# CRI:PAN
# HND:NIC
# HND:SLV

# MEX:USA
# MMR:THA
# MYS:IDN
# NIC:CRI
# SAU:KWT
# THA:MMR
# THA:MYS

farms_sf_correct <- farms_sf %>%
  mutate(GID_0 = as.character(GID_0)) %>%
  mutate(NAME_0 = as.character(NAME_0)) %>%
  mutate(NAME_1 = as.character(NAME_1)) %>%
  mutate(GID_0 = ifelse(check %in% "CHN:HKG", "CHN", GID_0)) %>%
  mutate(NAME_0 = ifelse(check %in% "CHN:HKG", "China", NAME_0)) %>%
  mutate(GID_0 = ifelse(check %in% "CHN:MAC", "CHN", GID_0)) %>%
  mutate(NAME_0 = ifelse(check %in% "CHN:MAC", "China", NAME_0)) %>%
  mutate(GID_0 = ifelse(check %in% "CRI:PAN", "CRI", GID_0)) %>%
  mutate(NAME_0 = ifelse(check %in% "CRI:PAN", "Costa Rica", NAME_0)) %>%  
  mutate(GID_0 = ifelse(check %in% "HND:NIC", "HND", GID_0)) %>%
  mutate(NAME_0 = ifelse(check %in% "HND:NIC", "Honduras", NAME_0)) %>%  
  mutate(GID_0 = ifelse(check %in% "HND:SLV", "HND", GID_0)) %>%
  mutate(NAME_0 = ifelse(check %in% "HND:SLV", "Honduras", NAME_0)) %>%   
  mutate(GID_0 = ifelse(check %in% "NIC:CRI", "NIC", GID_0)) %>%
  mutate(NAME_0 = ifelse(check %in% "NIC:CRI", "Nicaragua", NAME_0)) %>%  
  mutate(GID_0 = ifelse(check %in% "MEX:USA", "MEX", GID_0)) %>%
  mutate(NAME_0 = ifelse(check %in% "MEX:USA", "Mexico", NAME_0)) %>%   
  mutate(GID_0 = ifelse(check %in% "MMR:THA", "MMR", GID_0)) %>%
  mutate(NAME_0 = ifelse(check %in% "MMR:THA", "Myanmar", NAME_0)) %>% 
  mutate(GID_0 = ifelse(check %in% "MYS:IDN", "MYS", GID_0)) %>%
  mutate(NAME_0 = ifelse(check %in% "MYS:IDN", "Malaysia", NAME_0)) %>%
  mutate(GID_0 = ifelse(check %in% "SAU:KWT", "SAU", GID_0)) %>%
  mutate(NAME_0 = ifelse(check %in% "SAU:KWT", "Saudi Arabia", NAME_0)) %>%  
  mutate(GID_0 = ifelse(check %in% "THA:MMR", "THA", GID_0)) %>%
  mutate(NAME_0 = ifelse(check %in% "THA:MMR", "Thailand", NAME_0)) %>%  
  mutate(GID_0 = ifelse(check %in% "THA:MYS", "THA", GID_0)) %>%
  mutate(NAME_0 = ifelse(check %in% "THA:MYS", "Thailand", NAME_0))

@@

# FAO salmon production data per country
```{r}

@@

farms_sf <- st_as_sf(farms_prod)
mapview(farms_sf)

@@

write_csv(shrimp_raster_df, "/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/aquaculture/marine/shrimp/shrimp_farm_updated.csv")

>>>>

# all_farms <- readRDS("data/all_modeled_marine_farms.rds")

all_farms <- read_csv("data/all_marine_aquaculture_farms_sources_final.csv") %>%
      st_as_sf(., 
           coords = c("X", "Y"),
           crs = crs(food_raster), 
           agr = "constant")

shrimp_farms <- all_farms %>%
  filter(species_group == "Shrimps and prawns")

@@

# CHN:HKG
# CHN:MAC
# CHN:PRK
# CRI:PAN
# HND:NIC
# HND:SLV
# IRN:IRQ
# MEX:USA
# MYS:BRN
# MYS:SGP
# PAN:CRI
# SAU:KWT
# SLV:GTM
# TWN:CHN

farms_sf_correct <- farms_sf %>%
  mutate(GID_0 = as.character(GID_0)) %>%
  mutate(NAME_0 = as.character(NAME_0)) %>%
  mutate(NAME_1 = as.character(NAME_1)) %>%
  mutate(GID_0 = ifelse(check %in% "CHN:HKG", "CHN", GID_0)) %>%
  mutate(NAME_0 = ifelse(check %in% "CHN:HKG", "China", NAME_0)) %>%
  mutate(GID_0 = ifelse(check %in% "CHN:MAC", "CHN", GID_0)) %>%
  mutate(NAME_0 = ifelse(check %in% "CHN:MAC", "China", NAME_0)) %>%
  mutate(GID_0 = ifelse(check %in% "CHN:PRK", "CHN", GID_0)) %>%
  mutate(NAME_0 = ifelse(check %in% "CHN:PRK", "China", NAME_0)) %>%
  mutate(GID_0 = ifelse(check %in% "CRI:PAN", "CRI", GID_0)) %>%
  mutate(NAME_0 = ifelse(check %in% "CRI:PAN", "Costa Rica", NAME_0)) %>%  
  mutate(GID_0 = ifelse(check %in% "HND:NIC", "HND", GID_0)) %>%
  mutate(NAME_0 = ifelse(check %in% "HND:NIC", "Honduras", NAME_0)) %>%  
  mutate(GID_0 = ifelse(check %in% "HND:SLV", "HND", GID_0)) %>%
  mutate(NAME_0 = ifelse(check %in% "HND:SLV", "Honduras", NAME_0)) %>%  
  mutate(GID_0 = ifelse(check %in% "IRN:IRQ", "IRN", GID_0)) %>%
  mutate(NAME_0 = ifelse(check %in% "IRN:IRQ", "Iran", NAME_0)) %>%  
  mutate(GID_0 = ifelse(check %in% "MEX:USA", "MEX", GID_0)) %>%
  mutate(NAME_0 = ifelse(check %in% "MEX:USA", "Mexico", NAME_0)) %>% 
  mutate(GID_0 = ifelse(check %in% "MYS:BRN", "MYS", GID_0)) %>%
  mutate(NAME_0 = ifelse(check %in% "MYS:BRN", "Malaysia", NAME_0)) %>%
  mutate(GID_0 = ifelse(check %in% "MYS:SGP", "MYS", GID_0)) %>%
  mutate(NAME_0 = ifelse(check %in% "MYS:SGP", "Malaysia", NAME_0)) %>%
  mutate(GID_0 = ifelse(check %in% "PAN:CRI", "PAN", GID_0)) %>%
  mutate(NAME_0 = ifelse(check %in% "PAN:CRI", "Panama", NAME_0)) %>%    
  mutate(GID_0 = ifelse(check %in% "SAU:KWT", "SAU", GID_0)) %>%
  mutate(NAME_0 = ifelse(check %in% "SAU:KWT", "Saudi Arabia", NAME_0)) %>%  
  mutate(GID_0 = ifelse(check %in% "SLV:GTM", "SLV", GID_0)) %>%
  mutate(NAME_0 = ifelse(check %in% "SLV:GTM", "El Salvador", NAME_0)) %>%  
  mutate(GID_0 = ifelse(check %in% "TWN:CHN", "TWN", GID_0)) %>%
  mutate(NAME_0 = ifelse(check %in% "TWN:CHN", "Taiwan", NAME_0))

@@

# FAO shrimp production data per country
```{r}

@@

farms_sf <- st_as_sf(farms_prod)
mapview(farms_sf, cex = "tonnes_per_farm")

@@

write_csv(shrimp_raster_df, "/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/aquaculture/marine/shrimp/updated/shrimp_farm_updated.csv")

### 2021-08-12 marine/last_step_rasterize/tuna_rasterize_production.Rmd
cleaned up scripts and READMEs

all_farms <- readRDS("data/all_modeled_marine_farms.rds")

tuna_farms <- all_farms %>%
  dplyr::filter(type == "tuna")

@@

farms_sf_prod <- st_as_sf(farm_production)
mapview(farms_sf_prod)

## write the production data so that we can use it for the nutrient layer. 
st_write(farms_sf_prod, "/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/aquaculture/marine/tuna/prod_data_by_cell_iso3c.shp", delete_dsn = TRUE)

@@

tuna_raster_check <- rasterize(farms_sf, food_raster, field = "tonnes_per_farm", fun=sum)
plot(log(tuna_raster_check+1))
cellStats(tuna_raster_check, stat = "sum", na.rm = TRUE) # 37072.56

@@

write_csv(tuna_raster_df, "/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/aquaculture/marine/tuna/tuna_farm_updated.csv")

>>>>

# all_farms <- readRDS("data/all_modeled_marine_farms_final.rds")

all_farms <- read_csv("data/all_marine_aquaculture_farms_sources_final.csv") %>%
      st_as_sf(., 
           coords = c("X", "Y"),
           crs = crs(food_raster), 
           agr = "constant")

tuna_farms <- all_farms %>%
  dplyr::filter(species_group == "Bluefin tuna")

@@

farms_sf_prod <- st_as_sf(farm_production)
mapview(farms_sf_prod, cex = "tonnes_per_farm")

## write the production data so that we can use it for the nutrient layer. 
st_write(farms_sf_prod, "/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/aquaculture/marine/tuna/updated/prod_data_by_cell_iso3c.shp", delete_dsn = TRUE)

@@

tuna_raster_check <- rasterize(farms_sf, food_raster, field = "tonnes_per_farm", fun=sum)
plot(log(tuna_raster_check+1))
# s <- raster::select(tuna_raster_check)
# plot(s)

cellStats(tuna_raster_check, stat = "sum", na.rm = TRUE) # 37072.56

@@

write_csv(tuna_raster_df, "/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/aquaculture/marine/tuna/updated/tuna_farm_updated.csv")

### 2021-08-12 marine/marine_fish_general/finfish_farms/STEP1_fao_finfish_production.Rmd
cleaned up scripts and READMEs

## write out marine fish taxa and production for Juliette
fao_finfish_taxa_prod <- fao %>%
  filter(species %in% finfish_list) %>%  
  filter(!is.na(value)) %>%
  arrange(-value, country) %>%
  filter(value>=500) %>%  # countries with less do not appear to be stable producers at this point
  group_by(country, iso3c, species, year) %>%
  dplyr::filter(year == 2017) %>%
  summarize(fao_tonnes_production = sum(value)) %>%
  arrange(country, iso3c, species, year)

@@

```{r}
## write out bivalve taxa and production for Juliette
fao_finfish_taxa_prod <- fao %>%
  filter(species %in% finfish_list) %>%  
  filter(!is.na(value)) %>%
  arrange(-value, country) %>%
  group_by(country, year) %>%
  summarise(value = sum(value)) %>%
  filter(value>=500) %>% # countries with less do not appear to be stable producers at this point
  filter(year == 2017)

>>>>

## write out marine fish taxa and production
fao_finfish_taxa_prod <- fao %>%
  filter(species %in% finfish_list) %>%  
  filter(!is.na(value)) %>%
  arrange(-value, country) %>%
  filter(value>=500) %>%  # countries with less do not appear to be stable producers at this point
  group_by(country, iso3c, species, year) %>%
  dplyr::filter(year == 2017) %>%
  summarize(fao_tonnes_production = sum(value)) %>%
  arrange(country, iso3c, species, year)

@@

```{r}
## explore out bivalve taxa and production
fao_finfish_taxa_prod <- fao %>%
  filter(species %in% finfish_list) %>%  
  filter(!is.na(value)) %>%
  arrange(-value, country) %>%
  group_by(country, year) %>%
  summarise(value = sum(value)) %>%
  filter(value>=500) %>% # countries with less do not appear to be stable producers at this point
  filter(year == 2017)

### 2021-08-12 marine/marine_fish_general/finfish_farms/STEP3_gapfill_finfish.Rmd
updated figs and stats with newly run farms
cleaned up scripts and READMEs

---
title: "Gapfilling the number of farms to add for marine_fish_general countries based on tonnes of production"
output: html_document
---

@@

>>>>

---
title: "Gapfilling the number of farms to add for marine_fish_general countries based on tonnes of production"
author: "Gage Clawson (NCEAS, UCSB, OHI)"
date: "12/14/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

This script takes the number of farms information collected in STEP2 and calculates a global tonnes of production per farm, and calculates the number of remaining farms needed to place in our placement model.

@@

write.csv(tonnes_per_farm, "output/finfish_comp_tonnes_per_farm.csv", row.names = FALSE)

### 2021-08-12 marine/salmon/salmon_farms/STEP1_fao_salmon_production.Rmd
cleaned up scripts and READMEs

# include only salmon
fao_salmon <- fao %>%
  filter(species %in% salmon_list) %>%  
  filter(!is.na(value)) %>%
  arrange(-value, country) %>%
  filter(value>=500) %>%  # countries with less do not appear to be stable producers at this point
  group_by(country, iso3c, year) %>%
  summarize(fao_tonnes_production = sum(value)) %>%
  arrange(country, iso3c, year)

@@

sum(explore$fao_tonnes_production) # 2753752
sum(explore$percent) # 100.01
```

>>>>

# include only salmon
fao_salmon <- fao %>%
  filter(species %in% salmon_list) %>%  
  filter(!is.na(value)) %>%
  arrange(-value, country) %>%
  filter(value>=500) %>%  # countries with less do not appear to be stable producers at this point
  group_by(country, iso3c, year) %>%
  summarize(fao_tonnes_production = sum(value)) %>%
  arrange(country, iso3c, year) %>%
  filter(year == 2017)

@@

sum(explore$fao_tonnes_production) # 2753752
sum(explore$percent) # 100.01

unique(explore$country) # - 16 countries 
```

```{r}
## explore out salmon taxa and production
fao_salmon_taxa_prod <- fao %>%
  filter(species %in% salmon_list) %>%  
  filter(!is.na(value)) %>%
  arrange(-value, country) %>%
  group_by(country, species, year) %>%
  summarise(value = sum(value)) %>%
  filter(value>=500) %>% # countries with less do not appear to be stable producers at this point
  filter(year == 2017)

sum(fao_salmon_taxa_prod$value) # 2753752 - adds 458 tonnes

unique(fao_salmon_taxa_prod$country) # 16 countries

test <- fao_salmon_taxa_prod %>%
  filter(str_detect(species, "almon"))

unique(test$country)

sum(test$value)

2753752 - 2569126
```

### 2021-08-12 marine/salmon/salmon_farms/STEP3_gapfill_salmon.Rmd
updated figs and stats with newly run farms
cleaned up scripts and READMEs

---
title: "Untitled"
output: html_document
---

@@

>>>>

---
title: "Estimate tonnage per farm and farms to place"
author: "Gage Clawson (UCSB, NCEAS, OHI)"
date: "10/09/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

In this script we estimate the average tonnage per known farm for 2017 and estimate the remaining number of farms to place for data-limited countries.

@@

write.csv(tonnes_per_farm, "output/salmon_comp_tonnes_per_farm.csv", row.names = FALSE)

### 2021-08-12 marine/shrimp/shrimp_farms/STEP1_fao_shrimp_production.Rmd
cleaned up scripts and READMEs

## Read in fao data and get a species list for salmon 
species_list <- read.csv(here("marine/STEP1_species_groups/int/marine_groups.csv"))  %>%
  distinct(species, aq_group)

@@

## write out shrimp taxa and production for Juliette
fao_shrimp_taxa_prod <- fao %>%
  filter(species %in% shrimp_list) %>%  
  filter(!is.na(value)) %>%
  arrange(-value, country) %>%
  filter(value>=500) %>%  # countries with less do not appear to be stable producers at this point
  group_by(country, iso3c, species, year) %>%
  dplyr::filter(year == 2017) %>%
  summarize(fao_tonnes_production = sum(value)) %>%
  arrange(country, iso3c, species, year)

@@

>>>>

## Read in fao data and get a species list for shrimp
species_list <- read.csv(here("marine/STEP1_species_groups/int/marine_groups.csv"))  %>%
  distinct(species, aq_group)

@@

## write out shrimp taxa and production
fao_shrimp_taxa_prod <- fao %>%
  filter(species %in% shrimp_list) %>%  
  filter(!is.na(value)) %>%
  arrange(-value, country) %>%
  filter(value>=500) %>%  # countries with less do not appear to be stable producers at this point
  group_by(country, iso3c, species, year) %>%
  dplyr::filter(year == 2017) %>%
  summarize(fao_tonnes_production = sum(value)) %>%
  arrange(country, iso3c, species, year)

taxa_group <- fao_shrimp_taxa_prod %>%
  group_by(species) %>%
  summarise(tonnes = sum(fao_tonnes_production)) %>%
  mutate(percent = tonnes/sum(tonnes))

@@

```{r}
## explore out shrimp taxa and production 
fao_shrimp_taxa_prod <- fao %>%
  filter(species %in% shrimp_list) %>%  
  filter(!is.na(value)) %>%
  arrange(-value, country) %>%
  group_by(country, year) %>%
  summarise(value = sum(value)) %>%
  filter(value>=500) %>% # countries with less do not appear to be stable producers at this point
  filter(year == 2017)

sum(fao_shrimp_taxa_prod$value) # 5095394 - adds ~1000 tonnes

unique(fao_shrimp_taxa_prod$country) # 35 countries 
```

### 2021-08-12 marine/tuna/tuna_farms/STEP1_fao_tuna_production.Rmd
cleaned up scripts and READMEs

## write out tuna taxa and production for Juliette
fao_tuna_taxa_prod <- fao %>%
  filter(species %in% tuna_list) %>%  
  filter(!is.na(value)) %>%
  arrange(-value, country) %>%
  filter(value>=500) %>%  # countries with less do not appear to be stable producers at this point
  group_by(country, iso3c, species, year) %>%
  dplyr::filter(year == 2017) %>%
  summarize(fao_tonnes_production = sum(value)) %>%
  arrange(country, iso3c, species, year)

@@

```{r}
## write out bivalve taxa and production for Juliette
fao_tuna_taxa_prod <- fao %>%
  filter(species %in% tuna_list) %>%  
  filter(!is.na(value)) %>%
  arrange(-value, country) %>%
  group_by(country, year) %>%
  summarise(value = sum(value)) %>%
  filter(value>=500) %>% # countries with less do not appear to be stable producers at this point
  filter(year == 2017)

>>>>

## write out tuna taxa and production
fao_tuna_taxa_prod <- fao %>%
  filter(species %in% tuna_list) %>%  
  filter(!is.na(value)) %>%
  arrange(-value, country) %>%
  filter(value>=500) %>%  # countries with less do not appear to be stable producers at this point
  group_by(country, iso3c, species, year) %>%
  dplyr::filter(year == 2017) %>%
  summarize(fao_tonnes_production = sum(value)) %>%
  arrange(country, iso3c, species, year)

@@

```{r}
## explore out bivalve taxa and production 
fao_tuna_taxa_prod <- fao %>%
  filter(species %in% tuna_list) %>%  
  filter(!is.na(value)) %>%
  arrange(-value, country) %>%
  group_by(country, year) %>%
  summarise(value = sum(value)) %>%
  filter(value>=500) %>% # countries with less do not appear to be stable producers at this point
  filter(year == 2017)

### 2021-08-12 marine/tuna/tuna_farms/STEP3_gapfill_tuna.Rmd
updated figs and stats with newly run farms
cleaned up scripts and READMEs

---
title: "Gapfilling the number of farms to add for tuna countries based on tonnes of production"
output: html_document
---

@@

>>>>

---
title: "Gapfilling the number of farms to add for tuna countries based on tonnes of production"
author: "Gage Clawson (NCEAS, UCSB, OHI)"
date: "12/14/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

This script takes the number of farms information collected in STEP2 and calculates a global tonnes of production per farm, and calculates the number of remaining farms needed to place in our placement model.

@@

write.csv(tonnes_per_farm, "output/tuna_comp_tonnes_per_farm.csv", row.names = FALSE)
